```{r, include=FALSE}
opts_chunk$set(fig.path="inst/assets/img/", fig.width=6, fig.height=4, comment=NA, warning=FALSE)
```

rbison
======

Wrapper to the USGS Bison API. 

### Info

See [here](http://bison.usgs.ornl.gov/services.html) for API docs for the BISON API.


### Quick start

#### Install rbison

```{r}
# install.packages("devtools")
# library(devtools)
# install_github("rbison", "ropensci")
library(rbison)
```

Notice that the function `bisonmap` automagically selects the map extent to plot for you, 
being one of the contiguous lower 48 states, or the lower 48 plus AK and HI, or a global map

#### If some or all points outside the US, a global map is drawn, and throws a warning. . You may want to make sure the occurrence lat/long coordinates are correct.
##### get data
```{r}
out <- bison(species = "Phocoenoides dalli dalli", count = 10)
```

##### inspect summary
```{r}
bison_data(out)
```

##### map occurrences
```{r}
bisonmap(out)
```


####  All points within the US (including AK and HI)
##### get data
```{r}
out <- bison(species="Bison bison", count=600)
```

##### inspect summary
```{r}
bison_data(out)
```

##### map occurrences
```{r}
bisonmap(out)
```

####  All points within the contiguous 48 states
##### get data
```{r}
out <- bison(species="Aquila chrysaetos", count=600)
```

##### inspect summary
```{r}
bison_data(out)
```

##### map occurrences
```{r}
bisonmap(out)
```


####  With any data returned from a `bison` call, you can choose to plot county or state level data
##### Counties - using last data call for Aquila 
```{r}
bisonmap(out, tomap="county")
```

##### States - using last data call for Aquila 
```{r}
bisonmap(out, tomap="state")
```


####  You can also query BISON via their SOLR interface
##### The taxa service searches for and gives back taxonomic names

```{r}
bison_solr_tax(query="bear", method='common_name')
```

You can also do fuzzy searches like this (which you can't do using the `bison` function)

```{r}
bison_solr_tax(query="*bear")[1:2]
```

And you can search by scientific name

```{r}
bison_solr_tax(query="helianthus", method="scientific_name")
```

##### The occurrence service searches by scientific names and gives back occurrence data similar to data given back by the `bison` function

Searching for data and looking at output

```{r}
out <- bison_solr_occ(scientific_name='"Ursus americanus"', state_code='"New Mexico"', rows=50, fl="occurrence_date,scientific_name")
bison_data(input=out)
```

Mapping the data

```{r}
out <- bison_solr_occ(scientific_name='"Ursus americanus"', rows=200)
bisonmap(out)
```