<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{An R Markdown Vignette made with knitr}
-->

rbison vignette - wrapper to the USGS BISON API
======

### About the package

XXXXXXX

### Info

See [here](http://bison.usgs.ornl.gov/services.html) for API docs for the BISON API.


### Quick start

#### Install rbison

```{r install, comment=NA, warning=FALSE}
# install.packages("devtools"); library(devtools); install_github("rbison", "ropensci")
library(rbison); library(sp); library(plyr); library(httr); library(rjson); library(gridExtra)
library(mapproj); library(ggplot2); library(maps); library(rgeos); library(rgdal)
```

Notice that the function `bisonmap` automagically selects the map extent to plot for you, 
being one of the contiguous lower 48 states, or the lower 48 plus AK and HI, or a global map

#### If some or all points outside the US, a global map is drawn, and throws a warning. . You may want to make sure the occurrence lat/long coordinates are correct.
##### get data
```{r one, fig.width=6, fig.height=4, comment=NA, warning=FALSE}
out <- bison(species = "Helianthus annuus", count = 10)
```

##### inspect summary
```{r two, fig.width=6, fig.height=4, comment=NA, warning=FALSE}
bison_data(out)
```

##### map occurrences
```{r}
bisonmap(out)
```


#### All points within the US (including AK and HI)
##### get data
```{r four, fig.width=6, fig.height=4, comment=NA, warning=FALSE}
out <- bison(species="Bison bison", count=600)
```

##### inspect summary
```{r five, fig.width=6, fig.height=4, comment=NA, warning=FALSE}
bison_data(out)
```

##### map occurrences
```{r six, fig.width=6, fig.height=4, comment=NA, warning=FALSE}
bisonmap(out)
```

####  All points within the contiguous 48 states
##### get data
```{r seven, fig.width=6, fig.height=4, comment=NA, warning=FALSE}
out <- bison(species="Aquila chrysaetos", count=600)
```

##### inspect summary
```{r eight, fig.width=6, fig.height=4, comment=NA, warning=FALSE}
bison_data(out)
```

##### map occurrences
```{r nine, fig.width=6, fig.height=4, comment=NA, warning=FALSE}
bisonmap(out)
```


####  With any data returned from a `bison` call, you can choose to plot county or state level data
##### Counties - using last data call for Aquila 
```{r ten, fig.width=6, fig.height=4, comment=NA, warning=FALSE}
bisonmap(out, tomap="county")
```

##### States - using last data call for Aquila 
```{r eleven, fig.width=6, fig.height=4, comment=NA, warning=FALSE}
bisonmap(out, tomap="state")
```