<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{An R Markdown Vignette made with knitr}
-->

rbison vignette - wrapper to the USGS BISON API
======

### About the package

`rbison` is an R package to search and retrieve data from the USGS BISON service. `rbison` wraps R code around the BISON API to allow you to talk to the BISON database from R. 

BISON has occurrence data for the US only. 

BISON is a node of the Global Biodiversity Information Facility (GBIF) - i.e., you can get data that's available in BISON via GBIF instead if you want. 

********************

### Info

See [here](http://bison.usgs.ornl.gov/services.html) for API docs for the BISON API.

********************

### Quick start

#### Install rbison

```{r install, comment=NA, warning=FALSE, message=FALSE}
# install.packages("devtools"); library(devtools); install_github("rbison", "ropensci")
library(rbison); library(sp); library(plyr); library(httr); library(rjson); library(gridExtra)
library(mapproj); library(ggplot2); library(maps); library(rgeos); library(rgdal)
```

Notice that the function `bisonmap` automagically selects the map extent to plot for you, 
being one of the contiguous lower 48 states, or the lower 48 plus AK and HI, or a global map

#### If some or all points outside the US, a global map is drawn, and throws a warning. . You may want to make sure the occurrence lat/long coordinates are correct.
##### get data
```{r one, fig.width=6, fig.height=4, comment=NA, warning=FALSE, cache=TRUE}
out <- bison(species = "Helianthus annuus", count = 10)
```

##### inspect summary
```{r two, fig.width=6, fig.height=4, comment=NA, warning=FALSE, cache=TRUE}
bison_data(out)
```

##### map occurrences
```{r, cache=TRUE}
bisonmap(out)
```

********************

#### All points within the US (including AK and HI)
##### get data
```{r four, fig.width=6, fig.height=4, comment=NA, warning=FALSE, cache=TRUE}
out <- bison(species="Bison bison", count=600)
```

##### inspect summary
```{r five, comment=NA, warning=FALSE, cache=TRUE}
bison_data(out)
```

##### map occurrences
```{r six, fig.width=6, fig.height=4, comment=NA, warning=FALSE, cache=TRUE}
bisonmap(out)
```

********************

####  All points within the contiguous 48 states
##### get data
```{r seven, fig.width=6, fig.height=4, comment=NA, warning=FALSE, cache=TRUE}
out <- bison(species="Aquila chrysaetos", count=600)
```

##### inspect summary
```{r eight, fig.width=6, fig.height=4, comment=NA, warning=FALSE, cache=TRUE}
bison_data(out)
```

##### map occurrences
```{r nine, fig.width=6, fig.height=4, comment=NA, warning=FALSE, cache=TRUE}
bisonmap(out)
```

********************

####  With any data returned from a `bison` call, you can choose to plot county or state level data
##### Counties - using last data call for Aquila 
```{r ten, fig.width=6, fig.height=4, comment=NA, warning=FALSE, cache=TRUE}
bisonmap(out, tomap="county")
```

##### States - using last data call for Aquila 
```{r eleven, fig.width=6, fig.height=4, comment=NA, warning=FALSE, cache=TRUE}
bisonmap(out, tomap="state")
```

********************

####  Constrain search with county IDs or bounding boxes

##### Constrain search to a certain county.

Check out [this site](http://www.epa.gov/enviro/html/codes/state.html) to get state and county fips codes. Fips codes are like so: First two digits are the state code - last three are the county code. For example the *06* in  06037 is the state of California, and the *037* is the Los Angeles county.

```{r countyfips, comment=NA, warning=FALSE, cache=TRUE}
out <- bison(species="Helianthus annuus", countyFips = 06037)

# Inspect summary
bison_data(out)

# By default, the query only returned 10 records
bison_data(out, "data_df")
```

Or specify county by its actual name - probably much easier. 

```{r countyname, comment=NA, warning=FALSE, cache=TRUE}
out <- bison(species="Helianthus annuus", county = "Los Angeles")

# Inspect summary
bison_data(out)

# By default, the query only returned 10 records
bison_data(out, "data_df")
```

`bison` will help you if you spell the name wrong, or use a partial name. The results are not printed below, but you would get a prompt asking you to pick between the two counties that start with *Los*.

```{r countyname_wrong, eval=FALSE}
bison(species="Helianthus annuus", county = "Los")
```

##### Constrain search to a amorphous area. 

Check out the Wikipedia page [here](http://en.wikipedia.org/wiki/Well-known_text) for an in depth look at the options, terminology, etc.

```{r wellknowntext, comment=NA, warning=FALSE, cache=TRUE}
out <- bison(species="Helianthus annuus", aoi = "POLYGON((-111.06360117772908 38.84001566645886,-110.80542246679359 39.37707771107983,-110.20117441992392 39.17722368276862,-110.20666758398464 38.90844075244811,-110.63513438085685 38.67724220095734,-111.06360117772908 38.84001566645886))")

# Inspect summary
bison_data(out)

# The data
bison_data(out, "data_df")
```

##### Constrain search to a certain aoibbox.

An aoibbox uses the format minx,miny,maxx,maxy.

```{r boundbox, comment=NA, warning=FALSE, cache=TRUE}
out <- bison(species="Helianthus annuus", aoibbox = '-120.31,35.81,-110.57,40.21')

# Inspect summary
bison_data(out)

# The data, by default, the query only returned 10 records
bison_data(out, "data_df")
```